### 2014-09-14 notes.Rmd
app engine, writeup website

# Summer Data Challenge

## Intro

The challenge is an open-ended contest to find new insights using our exclusively prepared datasets, and propose how these new insights can be used to create social or economic value in the form of new products, services, or businesses.

Submissions will be judged on technical ability, creativity, practicality, and their potential for generating social or economic impact. Anybody can download our datasets and enter the contest and submissions will be judged by a panel made up of experts from academia and industry. The best submissions will be showcased on Imperial College's Data Science Institute website.

## London house prices datasets

1. Map of house sales over 5 year period. Solar theme, white flashes for house sale, brightness / persistence proportional to cost. js plotting library. Start off with ggplot2 svg then animate via d3? (this is icing on the cake really, not relevant for new insights)

2. Actionable insights. Relate percentages changes to other variables, as supplied by other datasets from e.g. guardian data. Machine learning, variable importance etc. What makes for a good local housing market?

## New product OR new service OR new business

Housing sale data, of interest to REITs, realtors, consumer housing market, individuals.

Product ideas:

* app for personal use. History of house sales in a given postcode, outlook for future growth, summary stats from other sources. Who'd use this? Eventually be bought out by Zoopla etc. for integration with their services. Do they offer this stuff already? 
  * Currently rightmove lists past house sales for an area
  * ONS offers clunky postcode datasets: http://www.neighbourhood.statistics.gov.uk/dissemination/LeadAreaSearch.do?a=7&i=1001&m=0&s=1410700363772&enc=1&areaSearchText=nw9+6ef&extendedList=false&searchAreas=
  * Crime statistics: http://www.crime-statistics.co.uk/postcode

Input postcode of interest, get nice visualisation of previous house sales, median rolling prices. Integrated with other data. Also offer surrouding postcode data for comparison.