### 2014-10-01 report/article_4.tex
logo swap

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Arsclassica Article
% LaTeX Template
% Version 1.1 (10/6/14)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Lorenzo Pantieri (http://www.lorenzopantieri.net) with extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
10pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
%twoside, % Two page layout (page indentation for binding and different headers)
headinclude,footinclude, % Extra spacing for the header and footer
BCOR5mm, % Binding correction
]{scrartcl}

\input{structure.tex} % Include the structure.tex file which specified the document structure and layout

\hyphenation{Fortran hy-phen-ation} % Specify custom hyphenation points in words with dashes where you would like hyphenation to occur, or alternatively, don't put any dashes in a word to stop hyphenation altogether

%% additional packages:
\usepackage{wrapfig}
\usepackage[left=1.2in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{caption}

%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR(S)
%----------------------------------------------------------------------------------------

\title{ \vspace{-3em}
\includegraphics[width=5cm]{Figures/datarea.pdf} \\
{\large \#SummerDataChallenge}} % The article title

\author{ {\large Benjamin L. Moore\textsuperscript{1}}} % The article author(s) - author affiliations need to be specified in the AUTHOR AFFILIATIONS block

\date{} % An optional date to appear under the author(s)

%----------------------------------------------------------------------------------------

% custom macros:

\newcommand*{\logo}{\includegraphics[scale=.22]{Figures/datarea.pdf}}

\begin{document}

%----------------------------------------------------------------------------------------
%	HEADERS
%----------------------------------------------------------------------------------------

\renewcommand{\sectionmark}[1]{\markright{\spacedlowsmallcaps{summerdatachallenge}}} % The header for all pages (oneside) or for even pages (twoside)
%\renewcommand{\subsectionmark}[1]{\markright{\thesubsection~#1}} % Uncomment when using the twoside option - this modifies the header on odd pages
\lehead{\mbox{\llap{\small\thepage\kern1em\color{halfgray} \vline}\color{halfgray}\hspace{0.5em}\rightmark\hfil}} % The header style

\pagestyle{scrheadings} % Enable the headers specified in this block

%----------------------------------------------------------------------------------------
%	TABLE OF CONTENTS & LISTS OF FIGURES AND TABLES
%----------------------------------------------------------------------------------------

\maketitle % Print the title/author/date block

\setcounter{tocdepth}{2} % Set the depth of the table of contents to show sections and subsections only

% \tableofcontents % Print the table of contents

% \listoffigures % Print the list of figures

% \listoftables % Print the list of tables

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------
\vspace{-4em}
\section*{Overview}

I analysed the {\bf London house prices} dataset is several innovative
ways. Firstly \emph{fractal context} shows how house sales within an
arbitrary area relate to those if neighbouring districs, up and down
the postcode hierarchy. Secondly, I employed \emph{ARIMA forecasting}
to extrapolate an area's price trend with an eye to suggesting
profitable investments. Finally, these and other metrics were combined
into an \emph{investment grade} for a given postcode area. Combined
this set of analyses forms the basis of \logo, a product aimed at
``democratising real estate investment'', offering individuals and
businesses quantitative real estate insights for data-driven investment
decisions.

A full report with additional visualisations is online at {\bf
  \href{http://blm.io/datarea}{blm.io/datarea} }and scripts to
reproduce all analyses are available from {\bf
  \href{http://github.com/blmoore/summerdatachallenge}{github.com/blmoore/summerdatachallenge}}.

%----------------------------------------------------------------------------------------
%	AUTHOR AFFILIATIONS
%----------------------------------------------------------------------------------------

{\let\thefootnote\relax\footnotetext{\textsuperscript{1} \textit{MRC
      Human Genetics Unit, University of Edinburgh, Scotland, United Kingdom}}}

%----------------------------------------------------------------------------------------

%\newpage % Start the article content on the second page, remove this if you have a longer abstract that goes onto the second page

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\section{Description of analyses}

When considering a property in a given area, you may ask a realtor
questions like: how do prices here compare to surrounding postcodes?
How have prices increased over time?  A real estate professional can
give an opinion based on their own limited sample size, but with a
large dataset we can address these questions quantitatively and
visualise the results.

\subsection{Fractal context}

\setlength{\intextsep}{0em}
\begin{wrapfigure}{R}{.45\textwidth}
\begin{center}
\includegraphics[width=.44\textwidth]{Figures/fractal.png}
\caption{ Fractal context of postcode SW18 4HU in Wandsworth, South London.}
\end{center}
\end{wrapfigure}

Letting and sales sites may currently list some recent sales, but it's
not currently possible to see a quantitative overview of property
prices in a region, within the context of its sector, district and
postcode.

These questions can be answered with a clear and intuitive data
visualisation of price distributions in the postcode hierarchy --- we
call this the fractal context of a property price (Figure 1).

\subsection{ARIMA modelling}

Given a time series of housing prices, it's of interest to investors
to speculate on potential investment returns for a given property
area. This can be done through time series analysis using the
autoregressive integrated moving average (ARIMA) model (Figure
2). Investment returns can be optimised by selecting areas with
maximal growth projections.\textsuperscript{$\star$} \\

\begin{figure}[h]
\begin{center}
\includegraphics[width=.9\textwidth]{Figures/arima.png}
\caption{Example ARIMA forecast of median house prices for a specific postcode.}
\end{center}
\end{figure}

\vspace{-2em}
\subsection{Investment grading}

Combining growth forecasts and contextual information, as well as
other derived metrics from this dataset, allow an approximate
investment grading, relative to other postcodes within the area
covered by this dataset. This intuitive output metric can help
prioritise areas ripe for property investment. \\

{\let\thefootnote\relax\footnotetext{\textsuperscript{$\star$}
    Data intended to assist investors and does not constitute
    investment advice; independent advice should be sought where appropriate. }}

%----------------------------------------------------------------------------------------
%	METHODS
%----------------------------------------------------------------------------------------
\vspace{-2em}
\section{Value generation}

\logo \hspace{.1em} would first be released as both a web service and
associated apps for mobiles and tablets. These would offer limited
analysis and visualisation for free, with user analytics and feedback
helping to validate the business model.

\begin{wrapfigure}{r}{.4\textwidth}
\begin{center}
\includegraphics[width=.39\textwidth]{Figures/mockup.png}
\caption{ iOS app mock-up.}
\end{center}
\end{wrapfigure}

A small data science team will gradually expand the range of analyses
offered, integrating novel public and privately-acquired datasets and
release them under a rolling subscription model, aimed at private
landlords and property speculators, as well as professional real
estate investment trusts (REITs). Significant revenue could be
generated through tailored partnerships with funds and high net-worth
individuals.

Additionally the business could pivot to an analytics provider for
existing letting and sales sites, offering unique data insights
indirectly to the consumer market, or as independent fund consultants
for the financial services. Having produced a first-to-market
innovative product positions \logo \hspace{.1em} to succeed as a
provider of advanced real estate investment analytics. \\

\begin{figure}[h!]
\begin{center}
\includegraphics[width=.1\textwidth]{Figures/logo.png} \\
\includegraphics[scale=.4]{Figures/datarea.pdf} \\
\caption*{ {\sf Democratising real estate investment}}
\end{center}
\end{figure}

\vspace{1em}
\begin{flushright}
{\small An enhanced overview of this entry is available online at: \href{http://blm.io/datarea}{blm.io/datarea}}
\end{flushright}

\end{document}