### 2014-10-09 report/article_4.tex
bringing web writeup inline with report

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Arsclassica Article
% LaTeX Template
% Version 1.1 (10/6/14)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Lorenzo Pantieri (http://www.lorenzopantieri.net) with extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
10pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
%twoside, % Two page layout (page indentation for binding and different headers)
headinclude,footinclude, % Extra spacing for the header and footer
BCOR5mm, % Binding correction
]{scrartcl}

\input{structure.tex} % Include the structure.tex file which specified the document structure and layout

\hyphenation{Fortran hy-phen-ation} % Specify custom hyphenation points in words with dashes where you would like hyphenation to occur, or alternatively, don't put any dashes in a word to stop hyphenation altogether

%% additional packages:
\usepackage{wrapfig}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{caption}

%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR(S)
%----------------------------------------------------------------------------------------

\title{ \vspace{-3em}
\includegraphics[width=.1\textwidth]{Figures/logo.png} \\
\includegraphics[width=4cm]{Figures/logotext.png} \\
} % The article title

\author{  Summer data challenge entry by Benjamin
  L. Moore\textsuperscript{1} } % The article author(s) - author affiliations need to be specified in the AUTHOR AFFILIATIONS block

\date{} % An optional date to appear under the author(s)

%----------------------------------------------------------------------------------------

% custom macros:

\newcommand*{\logo}{\includegraphics[scale=.04]{Figures/logotext.png}}

\begin{document}

%----------------------------------------------------------------------------------------
%	HEADERS
%----------------------------------------------------------------------------------------

%\renewcommand{\sectionmark}[1]{
\markright{\spacedlowsmallcaps{summerdatachallenge}}
%} % The header for all pages (oneside) or for even pages (twoside)
%\renewcommand{\subsectionmark}[1]{\markright{\thesubsection~#1}} % Uncomment when using the twoside option - this modifies the header on odd pages
\lehead{\mbox{\llap{\small\thepage\kern1em\color{halfgray} \vline}\color{halfgray}\hspace{0.5em}\rightmark\hfil}} % The header style

\pagestyle{scrheadings} % Enable the headers specified in this block

%----------------------------------------------------------------------------------------
%	TABLE OF CONTENTS & LISTS OF FIGURES AND TABLES
%----------------------------------------------------------------------------------------

% \maketitle % Print the title/author/date block

\setcounter{tocdepth}{2} % Set the depth of the table of contents to show sections and subsections only

% \tableofcontents % Print the table of contents

% \listoffigures % Print the list of figures

% \listoftables % Print the list of tables

% Make header in inkscape?
\begin{figure}[width=\textwidth]
\vspace{-3em}
\centering
\includegraphics[width=.7\textwidth]{Figures/header2.pdf}
\vspace{.5em}
\end{figure}

%suppres headers
\thispagestyle{plain}

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------
\vspace{-4em}
\section*{Overview}

\setlength{\intextsep}{0em}
\begin{wrapfigure}{R}{.3\textwidth}
%\vspace{-1em}
\centering
\includegraphics[width=.28\textwidth]{Figures/overview.png}
\vspace{-1em}
\caption*{ A map of the house prices dataset (brightness is proportional to
  price and density).}
\end{wrapfigure}

I analysed the {\bf London house prices} dataset is several innovative
ways. Firstly \emph{fractal context} shows how house sales within an
arbitrary area relate to those if neighbouring districs, up and down
the postcode hierarchy. Secondly, I fit \emph{ARIMA models} to
forecast an area's price trend into the future, with an eye to
suggesting profitable investments. Finally, these and other metrics
were combined into an \emph{investment grade} for a given postcode
area. Combined this set of analyses forms the basis of \logo, a
startup aimed at ``democratising real estate investment'', offering
individuals and businesses quantitative real estate insights for
data-driven investment decisions.

A full report with additional visualisations is online at {\bf
  \leavevmode\href{http://blm.io/datarea}{blm.io/datarea} }and scripts to
reproduce all analyses are available from {\bf
  \leavevmode\href{http://github.com/blmoore/summerdatachallenge}{github.com/blmoore/summerdatachallenge}}.

%----------------------------------------------------------------------------------------
%	AUTHOR AFFILIATIONS
%----------------------------------------------------------------------------------------

{\let\thefootnote\relax\footnotetext{\textsuperscript{1} \textit{MRC
      Human Genetics Unit, University of Edinburgh, Scotland, United Kingdom}}}

%----------------------------------------------------------------------------------------

%\newpage % Start the article content on the second page, remove this if you have a longer abstract that goes onto the second page

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\vspace{-1em}
\section*{Analyses}

When considering a property in a given area, you may ask a realtor
questions like: how do prices here compare to surrounding postcodes?
How have prices increased over time? And how have prices changed over
time? A real estate professional can give an opinion based on their
own limited sample size, but with a large dataset we can address these
questions quantitatively and visualise the results.

\subsection*{Fractal context}

Letting and sales sites may currently list some recent sales, but it's
not currently possible to see a

%\setlength{\intextsep}{0em}
\begin{wrapfigure}[16]{l}{.45\textwidth}
\vspace{.5em}
\centering
\includegraphics[width=.43\textwidth]{Figures/fractal.png}
\caption{ Fractal context of postcode SW18 4HU in Wandsworth, South London.}
%\vspace{35em}
\end{wrapfigure}

\noindent quantitative overview of property
prices in a region, within the context of its sector, district and
postcode.

These questions can be answered with a clear and intuitive data
visualisation of price distributions in the postcode hierarchy --- we
call this the fractal context of a property price (Figure 1).

A specific postcode or area can be marked within its distribution,
giving a clear view of a properties relative pricing which cannot be
gleaned just by browsing recent sales. The example view (Figure 1)
combines ranked violin density plots and a geo-heatmap for a clear,
intuitive view of a locations price context.

\subsection*{ARIMA modelling}

Given a time series of housing prices, it's of interest to investors
to speculate on potential investment returns for a given property
area. This can be done through time series analysis using the
autoregressive integrated moving average (ARIMA) model (Figure 2). The
model fitting procedure included regularisation that allowed a $drift$
term to model non-stationary trends, as well as coefficients capturing
periodicity or seasonal effects. Potential investment returns could be
optimised by selecting areas with
maximal growth projections. \\

\begin{figure}[h]
\begin{center}
\includegraphics[width=.9\textwidth]{Figures/arima.png}
\caption{Example ARIMA forecast of median house prices for a specific postcode.}
\end{center}
\end{figure}

\vspace{-1em}
\subsection*{Investment grading}

Combining growth forecasts and other derived metrics from this
dataset allow an approximate investment grading, relative to other
postcodes within the area covered by this dataset. This intuitive
output metric can help prioritise areas ripe for property investment.

As an example, Figure 3 shows the top 5 best-ranked postcode sectors
in which to invest, according to their 12-month price growth forecast
and historical annualised volatility --- thus combining the empirical
data with theoretical model outputs. These include South London
suburbs of Upper Tooting and Morden, Walthamstow, as well as Brockley,
whose train station within SE4 1 was linked with the London overground network in 2010.\\

\begin{figure}[h]
\begin{center}
\includegraphics[width=.9\textwidth]{Figures/t5.pdf}
\caption{Top 5 property investment sectors by projected returns and historical volatility.}
\end{center}
\end{figure}

The rankings used in Figure 3 made use of the $drift$ parameter of
ARIMA models, combined with statistical volatility over the five year
house prices dataset. Both parameters were converted to quantiles and
weighted equally; those sectors with the highest projected growth
and least historical volatility receive the highest investment grade
(on a scale of AAA, AA, A, BBB \ldots C).\textsuperscript{$\star$}

Returning to our original example, we can conclude that SW18 4HU is a
promising choice of area for property investment. It's approximately
median-priced for the sector, which is the cheapest of SW18 (in turn
part of the desirable SW London area; Figure 1). It has good growth
projections of XX in the next 12 months (Figure 2), and when combined
with volatility and compared with all $\approx2600$ sectors in the
dataset, receives the top AAA investment grade, placing it in the
90\textsuperscript{th} percentile for predicted price growth and low
historical volatility.

{\let\thefootnote\relax\footnotetext{\textsuperscript{$\star$}
    Data intended to assist investors and does not constitute
    investment advice; independent advice should be sought where appropriate. }}

%----------------------------------------------------------------------------------------
%	METHODS
%----------------------------------------------------------------------------------------
\clearpage
\section*{Value generation}

The above-described analyses form the basis of \logo\hspace{.1em}, a
start-up aimed at ``democratising real estate investment'' by making
deep-dive analytics data available both to a consumer and professional
investment market. Initially \logo\hspace{.1em} is a provider of
information aimed at investors but cannot dispense investment advice
until authorised to do so by the FCA. Three initial areas of interest
to commercialise and generate social and economic value are listed:

\subsection*{1. Analytics provider for online house sales and lettings
agents}

\logo\hspace{.1em} can provide {\bf real-time property area analytics} to
existing market leaders in {\bf online sales and lettings}.

\begin{wrapfigure}{l}{.25\textwidth}
\begin{center}
\includegraphics[width=.18\textwidth]{Figures/zoopla.png}
\includegraphics[width=.23\textwidth]{Figures/rightmove.jpg}
\caption*{ Potential partners. }
\end{center}
\end{wrapfigure}

There are currently no mainsteam property sales or rental sites that
attempts to provide any quantitative insights to contextualise a
property's list price. Some list recent nearby sales in plain text,
but the wealth of historical sales data available allows for much
richer displays (as demonstrated with the fractal context view,
\emph{above}), which offer valuable and practical insights to
prospective homebuyers.

This added feature would help differentiate a given sales and letting
site as well as increase the brand positioning of \logo\hspace{.1em}
within the real estate sector. Contextual maps could be drawn as
interactive D3.js visualisations and neatly integrated with a typical
property search. Additionally, web analytics data from our partnered
site could be fed-back to help further enhancement and expansion of
the data-driven insights and visualisations we offer.

\subsection*{2. Direct-to-consumer demo application}
\logo \hspace{.1em} would release both a {\bf web service} and associated
{\bf mobiles and tablet applications}. These would offer limited analysis and
visualisation without any fee, with collected user data and feedback helping to
validate the business model and develop our software.

\begin{wrapfigure}[12]{r}{.4\textwidth}
\centering
\includegraphics[width=.39\textwidth]{Figures/mockup.png}
\caption{ iOS app mock-up.}
%\vspace{-1em}
\end{wrapfigure}

Releasing these products aligns with our aims of democratising real
estate investment. Amateur property investors and prospective
landlords can make data-driven decisions, the likes of which are likely
already employed by large investment trusts and funds.

A small data science team will gradually expand the range of analyses
offered, integrating novel public and privately-acquired datasets and
release them under a rolling subscription model, aimed at private
landlords and property speculators, as well as professional real
estate investment trusts (REITs). Significant revenue could be
generated through tailored partnerships with funds and high net-worth
individuals.

\subsection*{3. Real estate data science consultancy}

Having established a property area analytics brand, \logo\hspace{.1em}
will look to develop business relationships with real estate
investment funds and high net-worth individuals as a {\bf quantitative
  property investment consultancy} and independent analyst for the
financial services. Our domain expertise and existing partnerships and
products would place us as a market leader in this undeveloped
segment.

% \begin{figure}[h!]
% \begin{center}
% \includegraphics[width=.1\textwidth]{Figures/logo.png} \\
% \includegraphics[scale=.12]{Figures/logotext.png} \\
% \caption*{ {\sf Democratising real estate investment}}
% \end{center}
% \end{figure}

\vspace{1em}
\begin{flushright} {\small An enhanced online version of this short
    report is available at:
    \leavevmode\href{http://blm.io/datarea}{blm.io/datarea}}
\end{flushright}

\end{document}