### 2014-11-07 report/sdc_report.tex
tidy, rm big files

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Arsclassica Article
% LaTeX Template
% Version 1.1 (10/6/14)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Lorenzo Pantieri (http://www.lorenzopantieri.net) with extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
10pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
%twoside, % Two page layout (page indentation for binding and different headers)
headinclude,footinclude, % Extra spacing for the header and footer
BCOR5mm, % Binding correction
]{scrartcl}

\input{structure.tex} % Include the structure.tex file which specified the document structure and layout

\hyphenation{Fortran hy-phen-ation} % Specify custom hyphenation points in words with dashes where you would like hyphenation to occur, or alternatively, don't put any dashes in a word to stop hyphenation altogether

%% additional packages:
\usepackage{wrapfig}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{caption}

%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR(S)
%----------------------------------------------------------------------------------------

\title{ \vspace{-3em}
\includegraphics[width=.1\textwidth]{Figures/logo.png} \\
\includegraphics[width=4cm]{Figures/logotext.png} \\
} % The article title

\author{  Summer data challenge entry by Benjamin
  L. Moore\textsuperscript{1} } % The article author(s) - author affiliations need to be specified in the AUTHOR AFFILIATIONS block

\date{} % An optional date to appear under the author(s)

%----------------------------------------------------------------------------------------

% custom macros:

\newcommand*{\logo}{\includegraphics[scale=.04]{Figures/logotext.png}}

\begin{document}

%----------------------------------------------------------------------------------------
%	HEADERS
%----------------------------------------------------------------------------------------

%\renewcommand{\sectionmark}[1]{
\markright{\spacedlowsmallcaps{summerdatachallenge}}
%} % The header for all pages (oneside) or for even pages (twoside)
%\renewcommand{\subsectionmark}[1]{\markright{\thesubsection~#1}} % Uncomment when using the twoside option - this modifies the header on odd pages
\lehead{\mbox{\llap{\small\thepage\kern1em\color{halfgray} \vline}\color{halfgray}\hspace{0.5em}\rightmark\hfil}} % The header style

\pagestyle{scrheadings} % Enable the headers specified in this block

%----------------------------------------------------------------------------------------
%	TABLE OF CONTENTS & LISTS OF FIGURES AND TABLES
%----------------------------------------------------------------------------------------

% \maketitle % Print the title/author/date block

\setcounter{tocdepth}{2} % Set the depth of the table of contents to show sections and subsections only

% \tableofcontents % Print the table of contents

% \listoffigures % Print the list of figures

% \listoftables % Print the list of tables

\begin{figure}[width=\textwidth]
\vspace{-3em}
\centering
\includegraphics[width=.7\textwidth]{Figures/header2.pdf}
\vspace{.5em}
\end{figure}

%suppres headers
\thispagestyle{plain}

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------
\vspace{-4em}
\section*{Overview}

\setlength{\intextsep}{0em}
\begin{wrapfigure}{R}{.3\textwidth}
%\vspace{-1em}
\centering
\includegraphics[width=.28\textwidth]{Figures/overview.png}
\vspace{-1em}
\caption*{ A map of the house prices dataset (brightness is proportional to
  price and density).}
\end{wrapfigure}

I analysed the {\bf London house prices} dataset in several creative
ways. Firstly, the \emph{fractal context} of an area visualises how house
prices relate to those of neighbouring areas at different scales up
and down the postcode hierarchy. Secondly, I fit \emph{ARIMA models}
to forecast an area's price trend into the future, with an eye to
suggesting profitable investments. Finally, growth and other metrics
were combined into an \emph{investment grade} for a given postcode
area. Together this set of analyses forms the basis of \logo, a
startup aimed at \emph{democratising real estate investment}, offering
individuals and businesses quantitative property area insights to
enable data-driven investment decisions.

An enhanced report with interactive visualisations is online at 
{\bf \leavevmode\href{http://blm.io/datarea}{blm.io/datarea} } 
and scripts to reproduce all analyses are available from 
{\bf
  \leavevmode\href{http://github.com/blmoore/summerdatachallenge}{github.com/blmoore/summerdatachallenge}
}.

%----------------------------------------------------------------------------------------
%	AUTHOR AFFILIATIONS
%----------------------------------------------------------------------------------------

%{\let\thefootnote\relax\footnotetext{\textsuperscript{1} \textit{MRC
 %     Human Genetics Unit, University of Edinburgh, Scotland, United Kingdom}}}

%----------------------------------------------------------------------------------------

%\newpage % Start the article content on the second page, remove this if you have a longer abstract that goes onto the second page

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\vspace{-1em}
\section*{Section A: Analyses}

When considering a property in a given area, you may want to ask
questions like: how do prices here compare to the surrounding areas?
How have they changed over time? Where will they be in a year's time?
A real estate professional can give an opinion based on their own
limited sample size, but with a large dataset we can address these
questions through statistical analysis and visualise the results.

\vspace{-.5em}
\subsection*{Fractal context}

Letting and sales sites sometimes list some recent sales in a given
area, but it's not currently possible to get a quantitative overview
of property prices in a region, within the context of its sector,
district and postcode.

%\setlength{\intextsep}{0em}
\begin{wrapfigure}[16]{l}{.45\textwidth}
%\vspace{.5em}
\centering
\includegraphics[width=.43\textwidth]{Figures/fractal.png}
\caption{ Fractal context of postcode SW18 4HU in Wandsworth, South London.}
%\vspace{35em}
\end{wrapfigure}

%\noindent

These questions can be answered with a clear and intuitive data
visualisation of price distributions in the postcode hierarchy --- what
I've called the fractal context of a property price (Figure 1).

Any chosen area can be seen in the context of its surroundings, from
full postcodes to much wider areas, clearly showing how the pricing of
any given property relates to those nearby.  The example graphic for
postcode SW18 4HU (Figure 1) combines ranked violin density plots and
a geo-heatmap for a clear, intuitive view of a location's price
context.

\vspace{-.5em}
\subsection*{ARIMA modelling}

Property investors need a method of evaluating the potential investment
returns of a given property area. This can be achieved through time
series modelling with autoregressive integrated moving average (ARIMA)
models (Figure 2). For this application regularised models were fit
which allowed a $drift$ term to capture non-stationary trends
(i.e. allowing for gain, loss or neither) as well as coefficients
reflecting periodicity or seasonal effects. Hence potential investment
returns could then be optimised by selecting areas with
maximal growth projections.

\begin{figure}[h]
\begin{center}
\includegraphics[width=.9\textwidth]{Figures/arima.png}
\caption{Example ARIMA forecast of median house prices for a specific postcode.}
\end{center}
\end{figure}

\vspace{-2em}
\subsection*{Investment grading}

Combining growth forecasts and with an additional measure of
variability allows an approximate investment grading of each sector,
relative to other postcodes within the area covered by this
dataset. This intuitive output metric can help prioritise areas ripe
for property investment and combines the theoretical model projections
with an empirical metric of past stability.

As an example, Figure 3 shows the top 5 postcode sectors in which to
invest, according to their 12-month price growth forecast and
historical annualised volatility. These include Brockley,
whose train station within SE4 1 was linked with the London Overground network in 2010.\\

\begin{figure}[h]
\begin{center}
\includegraphics[width=.9\textwidth]{Figures/t5.pdf}
\caption{Top 5 property investment sectors by projected returns and historical volatility.}
\end{center}
\end{figure}

\setlength{\intextsep}{0em}
\begin{wrapfigure}{R}{.45\textwidth}
\vspace{.5em}
\centering
\includegraphics[width=.44\textwidth]{inkscape/grading_2.png}
\vspace{-1em}
\caption{ Quantile transformation for investment grading
  (AAA rated sectors in \emph{navy}).}
\vspace{.5em}
\end{wrapfigure}

Growth forecasts and volatility were converted to quantiles and
equally weighted (Figure 4); those sectors with the highest projected growth
and least historical volatility receive the best investment grade
(on a scale of AAA, AA, A, BBB \ldots C).

Returning to our original example, SW18 4HU is a promising choice of
area for property investment.\textsuperscript{$\star$} It's
median-priced for the sector, which is the cheapest of SW18 (in turn
part of the desirable SW London area; Figure 1). It has good growth
projections of over £28,000 in the next year (Figure 2), and receives
the top AAA investment grade for being in the 90\textsuperscript{th}
percentile for predicted price growth and low historical volatility
compared to all other sectors covered by the dataset.

{\let\thefootnote\relax\footnotetext{\textsuperscript{$\star$}
    Data intended to assist investors but does not constitute
    investment advice; independent investment advice should be sought where
    appropriate. The value of investments can go down as well as up. }}

%----------------------------------------------------------------------------------------
%	METHODS
%----------------------------------------------------------------------------------------
\clearpage
\section*{Section B: Value generation}

The above-described analyses form the basis of \logo\hspace{.1em}, a
start-up aimed at ``democratising real estate investment'' by making
deep-dive analytics data available both to a consumer and professional
investment market. Initially \logo\hspace{.1em} is a provider of
information aimed at investors but cannot dispense investment advice
until authorised to do so by the FCA. Three initial areas of interest
to commercialise and generate social and economic value are listed:

\subsection*{1. Analytics provider for online house sales and lettings
agents}

\logo\hspace{.1em} can provide {\bf real-time property area analytics} to
existing market leaders in {\bf online sales and lettings}.

\begin{wrapfigure}{l}{.25\textwidth}
\begin{center}
\includegraphics[width=.18\textwidth]{Figures/zoopla.png}
\includegraphics[width=.23\textwidth]{Figures/rightmove.jpg}
\caption*{ Potential partners. }
\end{center}
\end{wrapfigure}

There are currently no mainsteam property sales or rental sites that
attempt to provide any quantitative insights that contextualise a
property's list price. Some list recent nearby sales in plain text,
but the wealth of historical sales data available allows for much
richer displays (as demonstrated with the fractal context view,
Section A: Figure 1, and by the interactive data-driven map shown in
Figure 5), which offer valuable and practical insights to prospective
homebuyers.

This added feature would help differentiate a given sales and letting
site as well as increase the brand positioning of \logo\hspace{.1em}
within the real estate sector. Contextual maps could be drawn as
interactive D3.js visualisations and neatly integrated with a typical
property search. Additionally, web analytics data from our partnered
site could be fed-back to help further enhancement and expansion of
the data-driven insights and visualisations we offer, such as how
often a given area is searched for on their site and by which
demographics.

\begin{figure}[h]
\vspace{1.5em}
\centering
\includegraphics[width=.85\textwidth]{Figures/gmapScreenshot.png}
\vspace{.5em}
\caption{ A basic embeddable web app built with the Google Maps API
  (viewable at \href{http://blm.io/datarea}{blm.io/datarea}). }
\end{figure}

\subsection*{2. Direct-to-consumer demo application}

\logo \hspace{.1em} will release an initial product both as a {\bf web
  service} and associated {\bf mobiles and tablet applications}
(Figure 6). These
offer limited analysis and visualisation without any fee, with
collected user data and feedback helping to validate the business
model and futher develop our software.

\begin{wrapfigure}[14]{r}{.46\textwidth}
\vspace{.5em}
\centering
\includegraphics[width=.42\textwidth]{Figures/mockup.png}
\caption{ iOS app mock-up.}
%\vspace{-1em}
\end{wrapfigure}

Releasing these products aligns with our aims of democratising real
estate investment. Amateur property investors and prospective
landlords can make data-driven decisions, the likes of which are 
already employed by analysts for large investment funds.

A small data science team will gradually expand the range of analyses
offered, integrating novel public and privately-acquired datasets and
release them to a premium app under a rolling subscription model,
aimed at private landlords and property speculators as well as
professional real estate investment trusts (REITs). Significant
revenue could be generated through tailored partnerships with funds
and high net-worth individuals.

\subsection*{3. Real estate data science consultancy}

Having established a property area analytics brand, \logo\hspace{.1em}
will look to develop business relationships with real estate
investment funds and high net-worth individuals as a {\bf quantitative
  property investment consultancy} and independent analyst for the
financial services. Our domain expertise and existing partnerships and
products would place us as a market leader in this largely undeveloped
segment.

\begin{wrapfigure}[10]{l}{.33\textwidth}
\vspace{5em}
\centering
\includegraphics[width=.2\textwidth]{Figures/logo.png}
\includegraphics[width=.2\textwidth]{Figures/logotext.png}
%\caption*{ {\sf Democratising real estate investment}}
\caption{ Branding. }
%\vspace{-4em}
\end{wrapfigure}

\section*{Summary}

In this report I've given a brief overview my attempts to elute
maximal insight from a single large dataset through statistical
modelling, visualisations and other data science techniques. Together
these form a narrative of investigative analysis into a single
exemplar postcode, and the assessment of the potential investment
returns within.

I have shown that these analyses have the potential to generate
valuable insights for a range of customer segments:

\setitemize[0]{leftmargin=4em,itemindent=2em}
\begin{itemize}

\item Online sales and letting agents would benefit from
plug-in analytics provided by \logo \hspace{.1em} --- offering their
users advanced insights into spatiotemporal property statistics.

\item A web and mobile app to make a set of analyses widely-available
  to amateur property investors, new homebuyers and landlords.

\item A consultancy specialising in real estate investment to partner
  with REITs and other specialist investment funds, as well as
  charitable trusts.
\end{itemize}

\vspace{2em}

\hrule

\vspace{1.5em}
%\begin{flushright}
\centering 
{\small An enhanced online version of this short
    report is available at:
    \leavevmode\href{http://blm.io/datarea}{blm.io/datarea}}
%\end{flushright}

%\vspace{2em}
% \begin{figure}[h!]
% \begin{center}
% \includegraphics[width=.15\textwidth]{Figures/logo.png} \\
% \includegraphics[scale=.11]{Figures/logotext.png} \\
% \caption*{ {\sf Democratising real estate investment}}
% \end{center}
% \end{figure}

\end{document}